Novel lattice Boltzmann method for simulation of strongly shear thinning viscoelastic fluids

### 2024-01-25 NovelLBM.tex
cellInShear

@@

	\todo{do}

@@

	Real-time deformability cytometry (RT-DC)\todo{cite} is a novel approach for single cell characterization. The cells are forced through a constriction. The observed deformation can be used to infer the mechanical properties of the cell. For this simulation we use periodic boundary conditions and a pressure gradient\todo{specify}. The resulting flow can be seen in figure \ref{fig:RT-DC_v}. The component typically associated with viscosity of the stress tensor can be see in figure \ref{fig:RT-DC_tau}. There are three qualitative differences compared to a purely viscous fluid. Directly after the inlet is a region of low stress. Due to the advection of the stress, the fluid here despite being inside the constriction has not reacted to it yet. This region therefore could produce unexpected cell behavior if one passes through it. Similarly directly after the constriction, the stress has not yet lowered. If one would divide the stress in these two regions by the strain rate in an attempt to obtain a viscosity one would find an extremely low and an extremely high viscosity respectively. Depending on the exact parameters \glqq viscosities\grqq{ }can be reached that cannot be obtained through rheological measurements with the fluid. Meaning these viscosities are not valid. This is even clearer along the walls directly after the outlet. Here the viscosites obtained in this manner become negative. This shows the breakdown of the viscosity interpretation in viscoelastic cases and signifies the regions of interesting behavior. This also shows why viscoelastic simulations cannot always be replaced by shear thinning models and should be considered despite the high computational cost.

>>>>

	%We use richard/vtk 4dbc4d379842c538f084e68a5dcfe2555a674495 (2024-01-09)

@@

	A cell with a radius of \SI{6}{\micro\meter} and Youngs Modulus of \SI{100}{\pascal} is placed in a shear flow with a shear rate of \SI{400}{\per\second}. The fluid is described by the parameters of \SI{0.49}{\percent} methyl cellulose solution mentioned above. The lattice with a resolution of \num{100} lattice points is chosen significantly bigger than the cell to assure negligible impact of the walls. The lattice is periodic in $x$ and $z$ direction. At the edges of the simulation box in $y$ direction Dirichlet boundary conditions are imposed. The influence of the cell placed in the center of the simulation volume can barely seen in the velocity field (see figure \ref{fig:cellInShear_v}). In the von Mieses stress seen in figure \ref{fig:cellInShear_vM} it is the dominant feature. One can clearly see the shape of the cell by the depression of the stress it causes. Notable from this shape outwards radiate section of high and low stress. These are particularly pronounced along the longest axis of the deformed cell. From these one can see that the cell disturbs the fluid far from itself.
		\adjincludegraphics[trim=0 0 0 0, clip, scale=1]{assets/plots/cellInShear_v.eps}
		\caption{Plot of the velocity field around a cell in shear flow.}
		\label{fig:cellInShear_v}
		\adjincludegraphics[trim=0 0 0 0, clip, scale=1]{assets/plots/cellInShear_sigma_vM.eps}
		\caption{Plot of the von Mises stress around a cell in shear flow.}
		\label{fig:cellInShear_vM}

@@

	Real-time deformability cytometry (RT-DC)\todo{cite} is a novel approach for single cell characterization. The cells are forced through a constriction. The observed deformation can be used to infer the mechanical properties of the cell. For this simulation we use periodic boundary conditions and a pressure gradient\todo{specify}. The resulting flow can be seen in figure \ref{fig:RT-DC_v}. The component typically associated with viscosity of the stress tensor can be seen in figure \ref{fig:RT-DC_tau}. There are three qualitative differences compared to a purely viscous fluid. Directly after the inlet is a region of low stress. Due to the advection of the stress, the fluid here despite being inside the constriction has not reacted to it yet. This region therefore could produce unexpected cell behavior if one passes through it. Similarly, directly after the constriction, the stress has not yet lowered. If one would divide the stress in these two regions by the strain rate in an attempt to obtain a viscosity one would find an extremely low and an extremely high viscosity respectively. Depending on the exact parameters \glqq viscosities\grqq{ }can be reached that cannot be obtained through rheological measurements with the fluid. Meaning these viscosities are not valid. This is even clearer along the walls directly after the outlet. Here the viscosites obtained in this manner become negative. This shows the breakdown of the viscosity interpretation in viscoelastic cases and signifies the regions of interesting behavior. This also shows why viscoelastic simulations cannot always be replaced by shear thinning models and should be considered despite the high computational cost.
